---
title: "Check mixfdr in low noise simulation"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

Here we check that mixfdr actually does OK at fitting the bimodal distribution
when noise is small compared with bimodality.

```{r chunk_options, include=FALSE}
# Specify settings for displaying the plots in the rendered document.
source("chunk-options.R")
```

Simulate bimodal data with well-separated peaks. The histogram shows the
multi-modal nature of the underlying g is clear.

```{r sim_data}
source("../code/dsc-shrink/datamakers/datamaker.R")
sim.bimodal =
  rnormmix_datamaker(args = list(g = normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
                                 min_pi0   = 0.4,
                                 max_pi0   = 0.4,
                                 nsamp     = 1000,
                                 betahatsd = 1))
hist(sim.bimodal$input$betahat,nclass = 20,xlim = c(-10,10),prob = TRUE,
     main="histogram of simulated betahat values")
```

Now run mixfdr on the simulated data. Note that the fit captures the
multi-modal nature of g, although mixfdr continues to overestimate pi0.

```{r fit_models}
source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
sim.bimodal.mixfdroutput =
  mixfdr.wrapper(sim.bimodal$input,args = list(theonull = TRUE))
g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
x = seq(-10,10,length = 100)
par(cex.main = 1,font.main = 1)
plot(ecdf(sim.bimodal$meta$beta),cex = 0.6,
     main = "CDF of true beta: empirical (black) and fitted (red)")
lines(x,mixcdf(g,x),lwd = 2,lty = "dotted",col = "orangered")
```

## Session information

```{r info}
sessionInfo()
```
