---
title: "Estimate Non-zero mode"
author: "Matthew Stephens"
date: 2016-02-01
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

The purpose of this analysis is to check the performance of the non-zero mode option.

## Simple simulation

I conjecture that the mean of the posterior means should be close to the optimum for the mode...
maybe even equal to it. (That is, this would constitute a fixed point for the update. We aren't
explicitly using that directly in the current implementation; the uniform mixture uses optim to
do it numerically; the normal mixture uses a true EM I think...)

```{r}

check_mode = function(betahat, sebetahat,mixcompdist){
    z.ash=ashr::ash(betahat,sebetahat,mixcompdist=mixcompdist,nonzeromode=TRUE)
    average.posteriormean = mean(z.ash$PosteriorMean)
    fitted.mode = ashr::comp_mean(z.ash$fitted.g)[1]
    
    z.ash1 = ashr::ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
    g = z.ash1$fitted.g
    loglik = z.ash1$loglik
    
    loglik.down = ashr::ash(z-fitted.mode-0.01,1,g=g)$loglik
    loglik.up  = ashr::ash(z-fitted.mode+0.01,1,g=g)$loglik
    loglik.posteriormean = ashr::ash(z-average.posteriormean,1,g=g)$loglik

    return(list(fitted.mode=fitted.mode, average.posteriormean =average.posteriormean,
         loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
  }
  set.seed(100)
  z = rnorm(1000)+3
  check_mode(z,1,mixcompdist="uniform")
  check_mode(z,1,mixcompdist="normal")
  check_mode(z,1,mixcompdist="halfuniform")
```

```{r}
  set.seed(100)
  beta = rexp(1000)
  betahat = beta + rnorm(1000,0,0.1)
  z.ash.hu = ashr::ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu = ashr::ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
  
  z.ash.hu2 = ashr::ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu2 = ashr::ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
  
  z.ash.hu.nzm = ashr::ash(betahat,0.1,mixcompdist="halfuniform",nonzeromode=TRUE,method="shrink")
  z.ash.hu$loglik
  z.ash.hu.nzm$loglik
  
  test.op = function(c){return(-ashr::ash(betahat-c,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")$loglik)}
  optimize(test.op,interval=c(min(betahat),max(betahat)))
#  check_mode(betahat,.1,mixcompdist="halfuniform")
```

Note to self: check the normal version works too; check nonzeromodeEMobj  as it doesn't seem needed.

## Session information

```{r info}
sessionInfo()
```

