---
title: "Investigate poor coverage for ash.hu"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

First, we load the necessary libraries, and specify settings for
displaying the plots in the rendered document.

```{r chunk_options, include=FALSE}
library(ashr)
library(reshape2)
library(ggplot2)
library(dplyr)
library(dscr)
source("chunk-options.R")
```

Compile the tables which we will use to examine the results of the simulation
experiments.

```{r compile_tables}
load("../output/dsc-shrink-files/res.RData")

neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars = c("method","scenario","seed",".id"),value.name = "negprob") %>%
    filter(negprob > 0.95)
 
poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars = c("method","scenario","seed",".id"),value.name = "posprob") %>%
    filter(posprob > 0.95)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars = c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

ash.hu.ft = reslong.pos %>% filter(method == "ash.hu.s" & 
                                   scenario == "flat-top")
```

Extract an example to illustrate coverage of ash.hu.

```{r show_example}
eg = load_example(dsc_shrink,42,"flat-top",methodname = "ash.hu.s",
                  homedir = "../code/dsc-shrink")
out        <- eg$output$fitted_g
class(out) <- "list"
out        <- as.data.frame(out)
print(subset(out,pi > 0.01))
```

Notice how almost all the inferred weight is on a small positive component. 
As a result, false sign rate will be small, and there will be a strong tendency
to overestimate zero effects. This leads to coverage problems observed.

Now let's look at an example with u and spiky which seems to be somewhat 
badly calibrated for the negative discoveries.

```{r show_more_examples}
ash.u.s.spiky = reslong.neg %>% filter(method == "ash.u.s" & 
                                       scenario == "spiky")
ash.n.s.spiky = reslong.neg %>% filter(method == "ash.n.s" & 
                                         scenario == "spiky")
hist(ash.u.s.spiky$value,nclass = 100,
     main = paste("histogram of quantile where observation falls in its CI"))
hist(ash.n.s.spiky$value,nclass = 100,
     main = paste("histogram of quantile where observation falls in its CI"))
```

So what seems to be happening here is that the uniform tail is too short; 
when observation falls outside of this tail it gets a zero quantile of 
posterior interval.

Can we find an example illustrating this trend? (Note: the next chunk below
did not work for me, so I se* `eval=FALSE, include=FALSE`. Perhaps this can be 
fixed at some point. -Peter)

```{r example_short_tail, eval=FALSE, include=FALSE}
table((ash.u.s.spiky %>% filter(value == 0))$seed)
eg = load_example(dsc_shrink,seed = 39,scenario = "spiky",
                  methodname = "ash.u.s","../code/dsc-shrink")
```

For comparison, here are the positive discoveries; here they are not too bad.

```{r histograms_pos}
ash.u.s.spiky = reslong.pos %>% filter(method == "ash.u.s" & scenario == "spiky")
ash.n.s.spiky = reslong.pos %>% filter(method == "ash.n.s" & scenario == "spiky")
hist(ash.n.s.spiky$value)
hist(ash.u.s.spiky$value)
```

## Session information

```{r info}
sessionInfo()
```
