---
title: "DDD plots"
author: "Matthew Stephens"
date: 2016-10-25
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

This file produces ``meta-analysis" style plots that illustrate the shrinkage behaviour of ash on some simple simulated data.

First, we load the necessary libraries.

```{r packages}
library(knitr)
library(rmeta)
library(ashr)
```

```{r chunk_options, include=FALSE}
# Specify settings for displaying the plots in the rendered document.
source("chunk-options.R")
opts_chunk$set(echo = TRUE)
```

## Simulate Data

We simulate 50 effects ($\beta$), half of which are truly "null" (ie 0) and 
the other half are $N(0,2^2)$. The observed effects ($\hat\beta$) all have 
standard error 1 here.

```{r chunk_label}
set.seed(1)
n        = 50
beta     = c(rep(0,n/2),rnorm(n/2,0,2))
betahat  = beta + rnorm(n)
beta.ash = ash(betahat,1)
```

## Meta Plots

Here we plot the standard Confidence Intervals, ordered by the observed effect 
sizes. The red colors indicate observations corresponding to effects that were 
truly 0.

```{r plot_unshrunk}
i = order(betahat)
metaplot(betahat[i],rep(1,n),ylab = "",xlab = "",xlim = c(-6,6),
         colors = meta.colors(box = c(rep("red",n/2),rep("black",n/2))[i]))
```

Here we plot the *shrunken* Confidence Intervals (more formally, posterior 
Credible Intervals), in the same order, with same color code. (We use the
approximation of the posterior mean $\pm$ 2 posterior standard deviation; 
really we should compute the posterior intervals using `ashci`, but this 
simple approximation suffices for illustrating the idea.)

Notice how the CIs are strongly shrunk towards 0, except for those that 
correspond to large effects, where the shrinkage is more modest. What is 
happening here is that ash recognizes that although many effects are null, 
others are quite big, so it avoids shrinking too much the intervals whose 
data are sufficiently strong to convince it that they are non-null.

```{r plot_shrunk}
metaplot(get_pm(beta.ash)[i],get_psd(beta.ash)[i],
         ylab = "",xlab = "",xlim = c(-6,6),
         colors = meta.colors(box = c(rep("red",n/2),rep("black",n/2))[i]))
```
