---
title: "Summarize distribution of LR statistic under null"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk_options, include=FALSE}
source("chunk-options.R")
```


```{r}
library(ggplot2)
library(dplyr)
load("../output/dsc-znull-files/res.znull.RData")
temp=ungroup(res %>% group_by(scenario,method) %>% summarise(gt0=mean(logLR>0)))
xtabs(gt0~method+scenario,temp)

```


Note that the ash.n.s conforms closely to the null expectation
of 0.5 chi2_0 +0.5 chi2_1 from Stram and Lee, Biometrics

conjecture: the uniform has the same asymptotic behaviour
and the half uniform is like the sum of two of these?

From this a 95% procedure would need to check for logLR > `r qchisq(0.9,df=1)/2`.

```{r}
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")

```

## Session information

```{r info}
sessionInfo()
```

